# Mass calculation for predicted metabolites
This skript makes use of different packages. *KEGGREST* is used to access the KEGG database. *RcppClassic* and *Rdisop* are used to calculate masses based on chemical formulas. *ggplot2* is used to plot a histogram of predicted masses. *purrr* is a functional programming toolkit used for data wrangling.

```{r required packages}
library(KEGGREST)
library(RcppClassic)
library(Rdisop)
library(ggplot2)
library(purrr)
```

# Calculate masses
Masses can be predicted for different occasions. The following shows different ways of creating a metabolite list, followed by the mass calculation for each dataset.

## Metabolite Mix
Solution of different Metabolites have been measured with MALDI-MS. The following shows two different ways of creating a list of metabolites. *A* import of a csv file, *B* creating the list in R.

```{r create list of MetaMix}
# read a .csv
MetaMix <- read.csv("MetaMix.csv",
                    header = TRUE,
                    sep = ",",
                    dec = ".")

# create dataframe in R
ID <-  c("C03557", "C00318", "C00185", "C00158", "C00475", "C04022", "C02679", "C00504", "C00031", "C00092", "C00037", "C00123", "C01384", "C00392", "C00140", "C01061", "C00079", "C00093", "C05682", "C00022", "C00121", "C00178", "C00086")
compound <- c("Aminomethylphosphonate", "Carnitine", "Cellobiose", "Citric acid" , "Cytidine", "Dimethylsulfoniopropionic acid", "Dodecanoic acid", "Folic acid", "Glucose", "Glucose-6-phosphate", "Glycine", "Leucine", "Maleic acid", "Mannitol", "N-Acetylglucosamine", "Nonanoic acid", "Phenylalanine", "Phosphoglyceric acid", "Phosphonoacetic acid", "Pyruvic acid", "Ribose", "Thymine", "Urea")

MetaMix <- data.frame(ID, compound, row.names = NULL)
MetaMix$ID <- as.character(MetaMix$ID)
MetaMix$compound <- as.character(MetaMix$compound)
```

Data acquisition from the KEGG database. Adds KEGG ID to the dataframe.

```{r query information of MetaMix from KEGG DB}
query_MetaMix <- list()
for (i in 1:nrow(MetaMix)){
  query_MetaMix[i] <- keggGet(MetaMix[i,
                                      "ID"])
}
```

Data acquisition from the KEGG database. Adds chemical formula and molecular weight to the dataframe.

```{r add information to MetaMix list}
for (i in 1:nrow(MetaMix)) {
  MetaMix$formula[i] <- query_MetaMix[[i]]$FORMULA
  MetaMix$mass[i]    <- getMolecule(MetaMix$formula[i])$exactmass
} # --> check chemical formular; no ions or polymers allowed
colnames(MetaMix) <- c("ID",
                       "compound",
                       "formula",
                       "mass")
```

Calculation of exact occuring masses based on chemical formula.

```{r add charges and adducts to MetaMix list}
MetaMix$pos_H    <- MetaMix$mass+getMolecule("H")$exactmass-0.00055
MetaMix$pos_Na   <- MetaMix$mass+getMolecule("Na")$exactmass-0.00055
MetaMix$pos_K    <- MetaMix$mass+getMolecule("K")$exactmass-0.00055
MetaMix$pos_DHB  <- MetaMix$mass+(getMolecule("C7H6O4")$exactmass-getMolecule("H2O")$exactmass)+getMolecule("H")$exactmass-0.00055
MetaMix$neg_H    <- MetaMix$mass-getMolecule("H")$exactmass+0.00055
MetaMix$neg_Cl   <- MetaMix$mass+getMolecule("Cl")$exactmass+0.00055

MetaMix[,5:ncol(MetaMix)] <- round(MetaMix[,5:ncol(MetaMix)], digit = 4)
```

## Hopanoids
```{r create list of Hopanoids}
# NIX metabolites
compound <- c("NAG-1", "NAG-2", "NAG-3", "NAG-4", "NAG-5", "Putrescine", "Alanine", "Glutamine", "Aspartate", "Adenine", "Adenosine", "AMP", "ADP", "ATP", "dAMP", "dADP", "dATP", "Guanine", "Guanosine", "GMP", "GDP", "GTP", "dGMP", "dGDP", "dGTP", "Cytosine", "Cytidine", "CMP", "CDP", "CTP", "dCMP", "dCDP", "dCTP", "Thymine", "Thymidine", "TMP", "TDP", "TTP", "dTMP", "dTDP", "dTTP")
formula <- c("C8H15NO6", "C16H28N2O11", "C24H41N3O16", "C32H54N4O21", "C40H67N5O26", "C4H12N2", "C3H7NO2", "C5H10N2O3", "C4H7NO4", "C5H5N5", "C10H13N5O4", "C10H14N5O7P", "C10H15N5O10P2", "C10H16N5O13P3", "C10H14N5O6P", "C10H15N5O9P2", "C10H16N5O12P3", "C5H5N5O", "C10H13N5O5", "C10H14N5O8P", "C10H15N5O11P2", "C10H16N5O14P3", "C10H14N5O7P", "C10H15N5O10P2", "C10H16N5O13P3", "C4H5N3O", "C9H13N3O5", "C9H14N3O8P", "C9H15N3O11P2", "C9H16N3O14P3", "C9H14N3O7P", "C9H15N3O10P2", "C9H16N3O13P3", "C5H6N2O2", "C10H14N2O5", "C10H15N2O9P", "C10H16N2O12P2", "C10H17N2O15P3", "C10H15N2O8P", "C10H16N2O11P2", "C10H17N2O14P3")



ID <-  c("C08627", "C21117", "C21118", "C21119", "C21120", "C21121", "C21122", "C21123")
compound <- c("Hopane-29-acetate", "Adenosylhopane", "Ribosylhopane", "Bacteriohopanetetrol" , "Bacteriohopanetetrol-acetylglucosamine", "Bacteriohopanetetrol-glucosamine", "Bacteriohopanetetrol cyclitol ether", "Aminobacteriohopanetriol")
Hopanoids <- data.frame(ID, compound, row.names = NULL)
Hopanoids$ID <- as.character(Hopanoids$ID)
Hopanoids$compound <- as.character(Hopanoids$compound)
```
```{r query information of Hopanoids from KEGG DB}
query_Hopanoids <- list()
for (i in 1:nrow(Hopanoids)){
  query_Hopanoids[i] <- keggGet(Hopanoids[i,
                                      "ID"])
}
```
```{r add information to Hopanoids list}
for (i in 1:nrow(Hopanoids)) {
  Hopanoids$formula[i] <- query_Hopanoids[[i]]$FORMULA
  Hopanoids$mass[i]    <- getMolecule(Hopanoids$formula[i])$exactmass
} # --> check chemical formular; no ions or polymers allowed
colnames(Hopanoids) <- c("ID",
                       "compound",
                       "formula",
                       "mass")
```
```{r add charges and adducts to Hopanoids list}
Hopanoids$pos_H    <- Hopanoids$mass+getMolecule("H")$exactmass-0.00055
Hopanoids$pos_Na   <- Hopanoids$mass+getMolecule("Na")$exactmass-0.00055
Hopanoids$pos_K    <- Hopanoids$mass+getMolecule("K")$exactmass-0.00055
Hopanoids$pos_DHB  <- Hopanoids$mass+getMolecule("C7H6O4")$exactmass-getMolecule("H2O")$exactmass+getMolecule("H")$exactmass-0.00055
Hopanoids$neg_H    <- Hopanoids$mass-getMolecule("H")$exactmass+0.00055
Hopanoids$neg_Cl   <- Hopanoids$mass+getMolecule("Cl")$exactmass+0.00055

Hopanoids[,5:ncol(Hopanoids)] <- round(Hopanoids[,5:ncol(Hopanoids)], digit = 4)
```

## KEGG
```{r list compounds of metabolic pathway}
listDatabases()
pathway                  <- keggList("pathway")
map                      <- keggGet("path:ko00680")
keggID                   <- as.data.frame(attributes(map[[1]]$COMPOUND))
compounds                <- as.data.frame(map[[1]]$COMPOUND)
KEGG_compounds           <- as.data.frame(c(keggID[1],
                                            compounds[1]))
colnames(KEGG_compounds) <- c("ID",
                              "compound")
```
```{r query information of pathway from KEGG DB}
query_KEGG_compounds <- list()
for (i in 1:nrow(KEGG_compounds)){
  query_KEGG_compounds[i] <- keggGet(KEGG_compounds[i,
                                                    "ID"])
}
```
```{r add information to KEGG list}
for (i in 1:nrow(KEGG_compounds)) {
  KEGG_compounds$formula[i] <- query_KEGG_compounds[[i]]$FORMULA
  KEGG_compounds$mass[i]    <- getMolecule(KEGG_compounds$formula[i])$exactmass
}
```
```{r add charges and adducts to KEGG list}
KEGG_compounds$pos_H    <- KEGG_compounds$mass+getMolecule("H")$exactmass-0.00055
KEGG_compounds$pos_Na   <- KEGG_compounds$mass+getMolecule("Na")$exactmass-0.00055
KEGG_compounds$pos_K    <- KEGG_compounds$mass+getMolecule("K")$exactmass-0.00055
KEGG_compounds$pos_DHB  <- KEGG_compounds$mass+getMolecule("C7H6O4")$exactmass-getMolecule("H2O")$exactmass+getMolecule("H")$exactmass-0.00055
KEGG_compounds$neg_H    <- KEGG_compounds$mass-getMolecule("H")$exactmass+0.00055
KEGG_compounds$neg_Cl   <- KEGG_compounds$mass+getMolecule("Cl")$exactmass+0.00055

KEGG_compounds[,5:ncol(KEGG_compounds)] <- round(KEGG_compounds[,5:ncol(KEGG_compounds)], digit = 4)
```

## PathwayTools data
### MOX
```{r import Pathway Tools smartTables of MOX}
MOX_compounds <- read.csv("smartTable/New_SmartTable_MOX.csv",
                          header = TRUE,
                          sep = ";",
                          dec = ".",
                          stringsAsFactors = TRUE)

```
```{r delete all lines without chemical formulas for MOX}
MOX_compounds <- subset(MOX_compounds,
                        MOX_compounds$KEGG != "")
MOX_compounds <- subset(MOX_compounds,
                        MOX_compounds$Chemical.Formula != "")
MOX_compounds <- MOX_compounds[, c("KEGG", "Name")]
colnames(MOX_compounds) <- c("ID",
                             "compound")
rownames(MOX_compounds) = NULL

# if compounds cannot be found: HTTP 404 -> manual subsetting
MOX_compounds <- rbind(MOX_compounds[1:203, ],
                       MOX_compounds[205:417, ],
                       MOX_compounds[419:531, ],
                       MOX_compounds[533:534, ],
                       MOX_compounds[536:547, ])
```
```{r query information of MOX from KEGG DB}
query_MOX <- list()
for (i in 1:nrow(MOX_compounds)){
  query_MOX[i] <- keggGet(MOX_compounds[i,
                                        "ID"])
}
```
```{r add information to MOX list}
for (i in 1:nrow(MOX_compounds)) {
  MOX_compounds$formula[i] <- query_MOX[[i]]$FORMULA
  MOX_compounds$mass[i]    <- getMolecule(MOX_compounds$formula[i])$exactmass
} # --> check chemical formula; no ions or polymers allowed
MOX_compounds <- subset(MOX_compounds,
                        mass > 60)
MOX_compounds <- subset(MOX_compounds,
                        mass < 900)
```
```{r calculate and add masses and adducts for MOX}
MOX_compounds$pos_H    <- MOX_compounds$mass+getMolecule("H")$exactmass-0.00055
MOX_compounds$pos_Na   <- MOX_compounds$mass+getMolecule("Na")$exactmass-0.00055
MOX_compounds$pos_K    <- MOX_compounds$mass+getMolecule("K")$exactmass-0.00055
MOX_compounds$pos_DHB  <- MOX_compounds$mass+getMolecule("C7H6O4")$exactmass-getMolecule("H2O")$exactmass+getMolecule("H")$exactmass-0.00055
MOX_compounds$neg_H    <- MOX_compounds$mass-getMolecule("H")$exactmass+0.00055
MOX_compounds$neg_Cl   <- MOX_compounds$mass+getMolecule("Cl")$exactmass+0.00055

MOX_compounds[,5:ncol(MOX_compounds)] <- round(MOX_compounds[,5:ncol(MOX_compounds)], digit = 4)
```

### NIX
```{r import Pathway Tools smartTables of NIX}
NIX_compounds <- read.csv("smartTable/New_SmartTable_NIX.csv",
                          header = TRUE,
                          sep = ";",
                          dec = ".",
                          stringsAsFactors = TRUE)

```
```{r delete all lines without chemical formulas for NIX}
NIX_compounds <- subset(NIX_compounds,
                        NIX_compounds$KEGG != "")
NIX_compounds <- NIX_compounds[, c("KEGG", "Name")]
colnames(NIX_compounds) <- c("ID",
                             "compound")
rownames(NIX_compounds) = NULL

# if compounds cannot be found: HTTP 404 -> manual subsetting
NIX_compounds <- rbind(NIX_compounds[1:16, ],
                       NIX_compounds[18:197, ],
                       NIX_compounds[199:206, ],
                       NIX_compounds[266:289, ],
                       NIX_compounds[304:322, ],
                       NIX_compounds[324:381, ],
                       NIX_compounds[383:485, ],
                       NIX_compounds[487:581, ],
                       NIX_compounds[583:584, ],
                       NIX_compounds[586:596, ])
```
```{r query information of NIX from KEGG DB}
query_NIX <- list()
for (i in 1:nrow(NIX_compounds)){
  query_NIX[i] <- keggGet(NIX_compounds[i,
                                        "ID"])
}
```
```{r add information to NIX list}
for (i in 1:nrow(NIX_compounds)) {
  NIX_compounds$formula[i] <- query_NIX[[i]]$FORMULA
  NIX_compounds$mass[i]    <- getMolecule(NIX_compounds$formula[i])$exactmass
} # --> check chemical formula; no ions or polymers allowed
NIX_compounds <- subset(NIX_compounds,
                        mass > 60)
NIX_compounds <- subset(NIX_compounds,
                        mass < 900)
```
```{r calculate and add masses and adducts for NIX}
NIX_compounds$pos_H    <- NIX_compounds$mass+getMolecule("H")$exactmass-0.00055
NIX_compounds$pos_Na   <- NIX_compounds$mass+getMolecule("Na")$exactmass-0.00055
NIX_compounds$pos_K    <- NIX_compounds$mass+getMolecule("K")$exactmass-0.00055
NIX_compounds$pos_DHB  <- NIX_compounds$mass+getMolecule("C7H6O4")$exactmass-getMolecule("H2O")$exactmass+getMolecule("H")$exactmass-0.00055
NIX_compounds$neg_H    <- NIX_compounds$mass-getMolecule("H")$exactmass+0.00055
NIX_compounds$neg_Cl   <- NIX_compounds$mass+getMolecule("Cl")$exactmass+0.00055

NIX_compounds[,5:ncol(NIX_compounds)] <- round(NIX_compounds[,5:ncol(NIX_compounds)], digit = 4)
```

### WF1
```{r import Pathway Tools smartTables of WF1}
WF1_compounds <- read.csv("smartTable/New_SmartTable_WF1.csv",
                          header = TRUE,
                          sep = ";",
                          dec = ".",
                          stringsAsFactors = TRUE)
```
```{r delete all lines without chemical formulas for WF1}
WF1_compounds <- subset(WF1_compounds,
                        WF1_compounds$KEGG != "")
WF1_compounds <- subset(WF1_compounds,
                        WF1_compounds$Chemical.Formula != "")
WF1_compounds <- WF1_compounds[, c("KEGG", "Name")]
colnames(WF1_compounds) <- c("ID",
                             "compound")
rownames(WF1_compounds) = NULL

# if compounds cannot be found: HTTP 404 -> manual subsetting
WF1_compounds <- rbind(WF1_compounds[1:247, ],
                       WF1_compounds[249:488, ],
                       WF1_compounds[490:624, ],
                       WF1_compounds[626:627, ],
                       WF1_compounds[629:639, ])
```
```{r query information of WF1 from KEGG DB}
query_WF1 <- list()
for (i in 1:nrow(WF1_compounds)){
  query_WF1[i] <- keggGet(WF1_compounds[i,
                                        "ID"])
}
```
```{r add information to WF1 list}
for (i in 1:nrow(WF1_compounds)) {
  WF1_compounds$formula[i] <- query_WF1[[i]]$FORMULA
  WF1_compounds$mass[i]    <- getMolecule(WF1_compounds$formula[i])$exactmass
}
WF1_compounds <- subset(WF1_compounds,
                        mass > 60)
WF1_compounds <- subset(WF1_compounds,
                        mass < 900)
```
```{r calculate and add masses and adducts for WF1}
WF1_compounds$pos_H    <- WF1_compounds$mass+getMolecule("H")$exactmass-0.00055
WF1_compounds$pos_Na   <- WF1_compounds$mass+getMolecule("Na")$exactmass-0.00055
WF1_compounds$pos_K    <- WF1_compounds$mass+getMolecule("K")$exactmass-0.00055
WF1_compounds$pos_DHB  <- WF1_compounds$mass+getMolecule("C7H6O4")$exactmass-getMolecule("H2O")$exactmass+getMolecule("H")$exactmass-0.00055
WF1_compounds$neg_H    <- WF1_compounds$mass-getMolecule("H")$exactmass+0.00055
WF1_compounds$neg_Cl   <- WF1_compounds$mass+getMolecule("Cl")$exactmass+0.00055

WF1_compounds[,5:ncol(WF1_compounds)] <- round(WF1_compounds[,5:ncol(WF1_compounds)], digit = 4)
```

# Export predicted metabolites as .csv
## MetaMix compounds
```{r export MetaMix list}
write.csv(MetaMix,
          file = "prediction_MetaMix.csv")
```

## Hopanoids compounds
```{r export Hopanoid list}
write.csv(Hopanoids,
          file = "prediction_Hopanoids.csv")
```

## KEGG compounds
```{r export KEGG list}
write.csv(KEGG_compounds,
          file = "prediction_KEGG.csv")
```

## NIX compounds
```{r export NIX list}
write.csv(NIX_compounds,
          file = "prediction_NIX.csv")
```

## MOX compounds
```{r export MOX list}
write.csv(MOX_compounds,
          file = "prediction_MOX.csv")
```

## WF1 compounds
```{r export WF1 list}
write.csv(WF1_compounds,
          file = "prediction_WF1.csv")
```


# combined datasets
```{r combine datasets}
combined_spots <- rbind(MetaMix[,1:4], WF1_compounds[,1:4])
combined_tissue <- rbind(MOX_compounds[,1:4], KEGG_compounds[,1:4], Hopanoids[,1:4])
combined_all <- rbind(combined_spots, combined_tissue)
```

# histogram
```{r plotting a histogram of predicted masses}
X <- combined_tissue[combined_tissue$mass < 1000, ]
ggplot(X,
       aes(x = mass)) +
  geom_histogram(binwidth = 2) +
  xlab('Exact Masses of Predicted Compounds') +
  ylab('Frequency') +
  theme(axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=15)) +
  theme(axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15)) 
ggsave('KEGG_pt_FreqChart<1000.png')
```
